---
title: "tidytuesdayw4"
author: "Zoe Sidana Bunnath"
format: html
---

## Introduction

Using this the tidytuesday data of this week, I explored the World Bank data on statistical performance of countries. The dataset cover 99 percent of the world population and extend from 2016-2023.

I want to visualize the data of the countries in the East Asia & Pacific region in order to explore the relationship between statistical performance and a country's income level or population in 2017 to 2023. 

## Load Libraries

```{r}
#| message: false
#| warning: false

library(tidytuesdayR) # for loading the tidy tuesday data
library(tidyverse) # for data wrangling + ggplot
library(here) # for saving files
library(viridis) # for color-blind friendly color pallete
library(plotly) # for interactive chart
library(htmlwidgets) #for saving the widget
library(gganimate) #for plot animation

```

## Load Data

```{r}
#| message: false
#| warning: false

#Load tidy tuesday data from November 25 2025

tuesdata <- tidytuesdayR::tt_load('2025-11-25') #load data of the week

spi_indicators <- tuesdata$spi_indicators
head(spi_indicators)

```

## Tidy the Data

```{r}
#| message: false
#| warning: false

eap_spi<-spi_indicators %>% 
  filter(year %in% 2017:2023, region == "East Asia & Pacific") %>% #only keep these rows
#only keep these columns
select(region, #Region
       country, #Country
       year, #Year of observation
       income, #Income level of country
       population, #Population of the country
       overall_score) #Overall statistical performance score

head(eap_spi)

```


## Plot the data

```{r}
#| message: false
#| warning: false
#| label: fig-eap-spi
#| fig-cap: "A Look at the relationship between statistical performance and income level of countries in the East-Asia and Pacific Region "

 
# Prepare the data for the interactive plot
eap_spi_anim <- eap_spi %>%
  mutate(
    pop = round(population / 1e6, 2),  # convert population to millions
    
    # Text that will show up when hovering over each bubble
    text = paste(
      "Country: ", country,
      "<br>Population (M): ", pop,
      "<br>Income level: ", income,
      "<br>Overall SPI score: ", overall_score,
      "<br>Year: ", year
    )
  )

# Create animated and interactive plot
p <- plot_ly(
  data = eap_spi_anim,
  x = ~overall_score, # x-axis: SPI score
  y = ~population,    # y-axis: population
  frame = ~year,  # animate by year
  type = "scatter",
  mode = "markers",
  size = ~pop,  # bubble size based on population
  sizes = c(20, 80), # control min and max bubble size
  color = ~country,
  text = ~text,
  hoverinfo = "text"
) %>%
  layout(
    title = "Statistical Performance and Income in East Asia and Pacific (2017–2023)",
    xaxis = list(title = "Overall statistical performance score"),
    yaxis = list(title = "Income level"),
    showlegend = FALSE  # turn off legend 
  ) %>%
  animation_opts(
    frame = 1000,   # how fast the animation moves
    easing = "linear",
    redraw = FALSE
  ) %>%
  animation_slider(
    currentvalue = list(prefix = "Year: ")
  )

p
```

## Save the plot
```{r}
#Save the widget
saveWidget(
  p, here("Tidy_Tuesday", "Output", "tidytuesdayw4.html")
)
```

## A New Things I Learned
With this Tidy Tuesday dataset, I learned how to animate a plot with the plotly package while keeping it interactive. I used the frame argument to show how the data changes over time, and at the same time I could still hover and zoom to explore the details. Watching the bubble plot transition by year helped me better understand patterns in population and statistical performance instead of only looking at a single year

